\noindent
Copyright (c) 2006   The Regents of the University of California.
Produced at the Lawrence Livermore National Laboratory.
Written by the HYPRE team, (hypre-users@llnl.gov), UCRL-CODE-222953.
All rights reserved.

\vspace{1em}\noindent
This notice is required to be provided under our contract with the U.S. 
Department of Energy (DOE). This work was produced at the University of 
California, Lawrence Livermore National Laboratory under Contract No. W7405-ENG-48 with the DOE.
 
\vspace{1em}\noindent
Neither the United States Government nor the University of California nor any 
of their employees, makes any warranty, express or implied, or assumes any 
liability or responsibility for the accuracy, completeness, or usefulness of any 
information, apparatus, product, or process disclosed, or represents that its use 
would not infringe privately-owned rights. 

\vspace{1em}\noindent
Also, reference herein to any specific commercial products, process, or 
services by trade name, trademark, manufacturer or otherwise does not 
necessarily constitute or imply its endorsement, recommendation, or favoring by 
the United States Government or the University of California. The views and 
opinions of authors expressed herein do not necessarily state or reflect those of 
the United States Government or the University of California, and shall not be 
used for advertising or product endorsement purposes. 

\vspace{1em}\noindent
This file is part of HYPRE (see http://www.llnl.gov/CASC/hypre/).
Please see the COPYRIGHT\_and\_LICENSE file for the copyright notice, 
disclaimer and the GNU Lesser General Public License.

\vspace{1em}\noindent
This program is free software; you can redistribute it and/or modify it
under the terms of the GNU General Public License (as published by the Free
Software Foundation) version 2.1 dated February 1999.
  
\vspace{1em}\noindent
This program is distributed in the hope that it will be useful, but WITHOUT
ANY WARRANTY; without even the IMPLIED WARRANTY OF MERCHANTABILITY or
FITNESS FOR A PARTICULAR PURPOSE.  See the terms and conditions of the
GNU General Public License for more details.

\vspace{1em}\noindent
You should have received a copy of the GNU Lesser General Public License
along with this program; if not, write to the Free Software Foundation,
Inc., 59 Temple Place, Suite 330, Boston, MA 02111-1307 USA
